Dynamic Witnesses for Static Type Errors

### 2015-11-20 paper/pldi16-submission/evaluation.tex
done with intro

\paragraph{Results}
\label{sec:results-complexity}
The results of the experiment are summarized in
Figure~\ref{fig:results-complexity}.
%
The average number of single-step reductions per trace is 26 for the
UCSD dataset (27 for the Seminal dataset) with a maximum of 2,744 (659
for Seminal). The average number of jumps per trace is 3 (also 3 for
Seminal) with a maximium of 177 (52 for Seminal). In both datasets at
least 90\% of the traces contain fewer than 10 jumps.
%

>>>>

\paragraph{Results}
\label{sec:results-complexity}
The results of the experiment are summarized in
Figure~\ref{fig:results-complexity}.
%
The average number of single-step reductions per trace is 26 for the
UCSD dataset (27 for the \uwbench\ dataset) with a maximum of 2,744 (659
for \uwbench). The average number of jumps per trace is 3 (also 3 for
\uwbench) with a maximium of 177 (52 for \uwbench). In both datasets at
least 90\% of the traces contain fewer than 10 jumps.
%

### 2015-11-20 paper/pldi16-submission/introduction.tex
done with intro

Obviously, the problem of finding witnesses is undecidable.
In fact, due to the necessarily conservative nature of static
typing, there may not even exist. Thus, our heuristic approach
is only useful if it can find \emph{compact} witnesses for
\emph{real-world} programs.

\paragraph{3. Evaluating Witnesses}
%
Our third contribution is an extensive evaluation of our approach
on two different sets of ill-typed programs obtained by instrumenting
compilers used in beginner's classes~(\S~\ref{sec:evaluation}).
%
The first is the \uwbench\ data set~\cite{lerner_searching_2007}
comprising \uwsize\ ill-typed programs.
%
The second is a new \ucsdbench\ data set, comprising \ucsdsize\
ill-typed programs.
%
We show that for both data sets, our technique is able to generate
witnesses for nearly 90\% of the programs, in under a second the
vast majority of cases.
%
Furthermore, we show that a simple interactive strategy yields
compact counterexample traces with fewer than 5 steps for 60\%
of the programs, and under 10 steps for 90\% of the programs.

\smallskip
In short, our results show that at least for beginners, ill-typed
programs \emph{do} go wrong, in the vast majority of cases, which,
in turn, opens the door to a novel dynamic way to explain, understand,
and appreciate the benefits of static typing.

>>>>

\paragraph{3. Evaluating Witnesses}
%
Of course, the problem of finding witnesses is
undecidable. In fact, due to the necessarily
conservative nature of static typing, there
may not even exist any witnesses for a given
ill-typed program.
%
Thus, our approach is a heuristic that is only useful
if it can find \emph{compact} witnesses for
\emph{real-world} programs.
%
Our third contribution is an extensive evaluation of our approach
on two different sets of ill-typed programs obtained by instrumenting
compilers used in beginner's classes~(\S~\ref{sec:evaluation}).
%
The first is the \uwbench\ data set~\cite{lerner_searching_2007}
comprising \uwsize\ ill-typed programs.
%
The second is a new \ucsdbench\ data set, comprising \ucsdsize\
ill-typed programs.
%
We show that for both data sets, our technique is able to generate
witnesses for nearly 90\% of the programs, in under a second the
vast majority of cases.
%
Furthermore, we show that a simple interactive strategy yields
compact counterexample traces with fewer than 5 steps for 60\%
of the programs, and under 10 steps for 90\% of the programs.

\smallskip
Our results show that in the vast majority of
cases, (novices') ill-typed programs \emph{do} go wrong.
This, in turn, opens the door to a novel dynamic way to
explain, understand, and appreciate the benefits of static typing.
